### Labs 7.8.2: Splines and 7.8.3: GAMs
## Objectives
1. Develop regression and smoothing splines to optimize model fit.
2. Interpret plots of spline fits.
3. Develop and explore Generalized Additive Models (GAMs) to study multiple features simultaneously.

Before we first fit Wage to Age using a regression spline, we create an age grid of all occurrences of ages in the dataset - 18 to 80.
```{r}
library(splines)
require(ISLR)
attach(Wage)

agelims=range(age)
age.grid=seq(from=agelims[1],to=agelims[2])

fit = lm(wage~bs(age,knots=c(25,40,60)),data=Wage)
pred=predict(fit,newdata=list(age=age.grid),se=T)
plot(age,wage,col="gray")
lines(age.grid,pred$fit,lwd=2)
lines(age.grid,pred$fit+2*pred$se,lty="dashed")
lines(age.grid,pred$fit-2*pred$se,lty="dashed")
```

Here we produced a spline with 3 knots - 6 degrees of freedom (or basis functions). We can also use `df`:

```{r}
dim(bs(age,knots=c(25,40,60)))
dim(bs(age,df=6))
attr(bs(age,df=6),"knots")  
```

Here, R chooses the 3 knots.

Now, in order to fit a natural spline, use ns():

```{r}
fit2=lm(wage~ns(age,df=4),data=Wage)
pred2=predict(fit2,newdata=list(age=age.grid),se=T)
plot(age,wage,col="gray")
lines(age.grid,pred2$fit,col="red",lwd=2)
```

In order to fit a smoothing spline, we use smooth.spline()

In the following piece of code, what differences do you expect in the plots of fit and fit2?

```{r}
plot(age,wage,xlim=agelims,cex=.5,col="darkgrey")
title("Smoothing Spline")
fit=smooth.spline(age,wage,df=16)
fit2=smooth.spline(age,wage,cv=TRUE)
fit2$df
lines(fit,col="red",lwd=2)
lines(fit2,col="blue",lwd=2)
legend("topright",legend=c("16 DF","6.8 DF"),
       col=c("red","blue"),lty=1,lwd=,cex=.8)
```


```{r}
plot(age,wage,xlim=agelims,cex=.5,col="darkgrey")
title("Local Regression")
fit=loess(wage~age,span=.2,data=Wage)
fit2=loess(wage~age,span=.5,data=Wage)
lines(age.grid,predict(fit,data.frame(age=age.grid)),
      col="red",lwd=2)
lines(age.grid,predict(fit2,data.frame(age=age.grid)),
      col="blue",lwd=2)
legend("topright",legend=c("Span=0.2","Span=0.5"),
       col=c("red","blue"),lty=1,lwd=2,cex=.8)
```

Here we have performed local linear regression using spans of 0.2 and 0.5: that is, each neighborhood consists of 20% or 50% of the observations.


We now fit a GAM to predict wage using natural spline functions of year and age. Education as a qualitative predictor.

This is basically a linear regression model using different basis functions, so lm() is appropriate.
```{r}
gam1=lm(wage~ns(year,4)+ns(age,5)+education,data=Wage)
```
How about if we want to use more general GAM components? Like smoothing splines?

```{r}
library(gam)
gam.m3=gam(wage~s(year,4)+s(age,5)+education,data=Wage)
par(mfrow=c(1,3))
plot(gam.m3,se=TRUE,col="blue")
plot.Gam(gam1,se=T,col="red")
```

All the terms are fit simultaneously, taking each other into account for the response.

In these plots, `year` looks linear. Perform ANOVA to study which model is the best fit: no year, linear year, or spline year.
```{r}
gam.m1=gam(wage~s(age,5)+education,data=Wage)
gam.m2=gam(wage~year+s(age,5)+education,data=Wage)
anova(gam.m1,gam.m2,gam.m3,test="F")
summary(gam.m3)
```

```{r}
preds = predict(gam.m2, newdata = Wage)
gam.lo = gam(wage~s(year,df=4)+lo(age,span=0.7)+education, data = Wage)
plot.Gam(gam.lo, se=TRUE, col="green")
```
  
We can also use lo() to create interactions before calling Gam.

```{r}
gam.lo.i = gam(wage~lo(year,age,span=0.5) + education, data=Wage)

library(akima)
plot(gam.lo.i)

gam.lr = gam(I(wage>250)~year + s(age,df=5)+education, family = binomial, data=Wage)
par(mfrow=c(1,3))
plot(gam.lr,se=T,col="green")
table(education, I(wage>250))
```

What do you notice above?

```{r}
gam.lr.s=gam(I(wage>250)∼year+s(age,df=5)+education,family=binomial,
             data=Wage,subset=(education!="1. < HS Grad"))
plot(gam.lr.s,se=T,col="purple")
```

