---
title: "HousePrices_Team4"
author: "Banjarian"
date: "4/21/2020"
output:
  html_document: default
  pdf_document: default
---

## Import libraries and the data

```{r}
library(dplyr)
library(gbm)
library(randomForest)
library(Metrics)
library(ggplot2)
library(corrplot)

train = read.csv('train.csv')
test = read.csv('test.csv')

```

## Check dimensions of trainset

```{r}
dim(train)
```

## Data Exploration

### Make new df called "all"
### Initially, we were preprocessing the train and test set separately. However, we ran into issues with data being processed differently, especially when imputing.


### Checking for NAs (not necessarily missing values)
```{r}
test$SalePrice <- NA
all <-  rbind(train,test)
names(which(sapply(all,anyNA)))
```


### Explore the response

```{r}
ggplot(data=all[!is.na(all$SalePrice),], aes(x=SalePrice)) +
        geom_histogram(fill="blue", binwidth = 10000) +
        scale_x_continuous(breaks= seq(0, 800000, by=100000))
```
```{r}
summary(all$SalePrice)
```

### Correlations
```{r}
numericFeatures <- which(sapply(all, is.numeric)) #index vector numeric variables
numericFeatureNames <- names(numericFeatures) #saving names vector 
print(numericFeatureNames)
cat('There are', length(numericFeatures)-1, 'numeric variables')
```

```{r}
all_numericFeatures <- all[, numericFeatures]
cor_numericFeatures <- cor(all_numericFeatures, use="pairwise.complete.obs")

cor_sorted <- as.matrix(sort(cor_numericFeatures[,'SalePrice'], decreasing = TRUE))

CorHigh <- names(which(apply(cor_sorted, 1, function(x) abs(x)>0.5))) # select a threshold for correlation
cor_numericFeatures <- cor_numericFeatures[CorHigh, CorHigh]

corrplot.mixed(cor_numericFeatures, tl.col="black", tl.pos = "lt")
```

## Deal with missing data


```{r}
all <- all %>% mutate_if(is.factor, funs(factor(replace(as.character(.), is.na(.), "NotApplicable"))))
all <- all %>% mutate_if(is.integer, funs(replace(as.integer(.),is.na(.), 0)))
```


## Imputing the missing data

### Numeric

```{r}
all$GarageYrBlt[is.na(all$GarageYrBlt)] <- mean(all$GarageYrBlt[!is.na(all$GarageYrBlt)])
all$GarageYrBlt <- as.integer((all$GarageYrBlt))

all$LotFrontage[is.na(all$LotFrontage)] <- 0
all$MasVnrArea[is.na(all$MasVnrArea)] <- 0
```

### Categorical
```{r}
all$MSZoning[all$MSZoning == 'NotApplicable'] <- names(sort(-table(all$MSZoning)))[1]
all$MSZoning <- as.factor(all$MSZoning)

all$Utilities[all$Utilities == 'NotApplicable'] <- names(sort(-table(all$Utilities)))[1]
all$Utilities <- as.factor(all$Utilities)

all$Exterior1st[all$Exterior1st == 'NotApplicable'] <- names(sort(-table(all$Exterior1st)))[1]
all$Exterior1st <- as.factor(all$Exterior1st)

all$Exterior2nd[all$Exterior2nd == 'NotApplicable'] <- names(sort(-table(all$Exterior2nd)))[1]
all$Exterior2nd <- as.factor(all$Exterior2nd)

all$MasVnrType[all$MasVnrType == 'NotApplicable'] <- names(sort(-table(all$MasVnrType)))[1]
all$MasVnrType <- as.factor(all$MasVnrType)

all$MasVnrType[all$MasVnrType == 'NotApplicable'] <- names(sort(-table(all$MasVnrType)))[1]
all$MasVnrType <- as.factor(all$MasVnrType)

all$KitchenQual[all$KitchenQual == 'NotApplicable'] <- names(sort(-table(all$KitchenQual)))[1]
all$KitchenQual <- as.factor(all$KitchenQual)

all$Functional[all$Functional == 'NotApplicable'] <- names(sort(-table(all$Functional)))[1]
all$Functional <- as.factor(all$Functional)

all$SaleType[all$SaleType == 'NotApplicable'] <- names(sort(-table(all$SaleType)))[1]
all$SaleType <- as.factor(all$SaleType)
```

# Modeling

### Linear Model
```{r}
myLM <- glm(SalePrice ~. -Id , data = all[1:1460,])
summary(myLM)

myLM.preds <- predict(myLM, all[1461:2919,]) 

```

### Bagging 

```{r}
bag.houses = randomForest(SalePrice ~.-Id ,data=all[1:1460,],
                          mtry=79, importance =TRUE)

yhat.bag = predict (bag.houses ,newdata =all[1461:2919,])

summary(yhat.bag)
```

### Creating a random forest
```{r}
rf.houses =randomForest(SalePrice ~.-Id ,data=all[1:1460,],
                        mtry=26, importance =TRUE)
yhat.rf = predict(rf.houses, newdata =all[1461:2919,])

importance(rf.houses) 
varImpPlot(rf.houses) # observing the relative impact of each feature
```

### Creating a boosting model
```{r}
library(gbm)
set.seed(4)
boost.houses=gbm(SalePrice ~.-Id ,data=all[1:1460,], distribution="gaussian",
                 n.trees=5000, interaction.depth = 4)
yhat.boost = predict(boost.houses, newdata =all[1461:2919,], n.trees=5000)
```

### Refining our boosting model
```{r}
boost.houses=gbm(SalePrice ~.-Id ,data=all[1:1460,], distribution="gaussian",
                 n.trees=15000, interaction.depth = 6)
yhat.boost2 = predict(boost.houses, newdata =all[1461:2919,], n.trees=15000)
```

### Preparing output for submission
```{r}
submission6 <- data.frame('Id' = as.integer(test$Id), 'SalePrice' = yhat.boost2)
# save our file
write.csv(submission6, 'submission6.csv',row.names = F)
```

